---
title: "Length Frequency Bubble Plot"
author: "Stock Assessment Group FWRI, Liz Herdter"
output: 
  html_document:
    toc: true
    toc_depth: 2
    toc_float: true
    number_sections: true
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE)

```

# _Main Objectives_
1. Produce bubble plots describing length frequencies.

This is the resource I used to create the plots is [here.](http://t-redactyl.io/blog/2016/02/creating-plots-in-r-using-ggplot2-part-6-weighted-scatterplots.html)


# _Start Here_ 
Load packages.
```{r warning=FALSE, message=FALSE}
library(tidyverse)
```

## Define User Inputs

The lowest length bin.
```{r}
lbinlow=2
```

The highest length bin.
```{r}
lbinhigh=90
```

How many breaks you want to have.
```{r}
lbinbreaks=8
```

Specify size range for min and max if you do not like the default base size ramp. 
```{r}
size_range_min = 1
size_range_max= 10
```

## Get Data
Here is some example data from the West Florida Shelf Hogfish Update.

```{r}
setwd("G:/DATA/SPECIES/HOGFISH/Hogfish_2017/AgeLength/Rexported_spreadsheets")
lengths = read.csv("WFS_LengthSS_Gear.csv", header=T)
comhl <- subset(lengths, Fleet==1) %>% select(Year, X2:X90) # subset for just 1 fleet
```

## Wrangle it the way the data need to be for ggplot

The data are not in the right format. Right now they are just a bunch of length observations in bins..
```{r}
head(comhl)
```

Gather the length bins (2-90) and give them a column name and arrange by year
```{r warning=FALSE}
com_lengths <- comhl %>% gather(X2:X90, key="Length_cat", value="N") %>% arrange(Year)
head(com_lengths)
```
The length_cat entries are in a weird format of X2-X90. So let's make a new column called Len_cat that only has the bin numbers. To do this remove the "x" with the gsub method. 
```{r warning=FALSE}
com_lengths <- com_lengths %>%
              mutate(Len_cat = as.numeric(gsub("X", "", Length_cat))) %>%
              select(-Length_cat)


head(com_lengths)
```
Last thing... Year needs to be a factor. 
```{r}
com_lengths$Year <- as.factor(com_lengths$Year)

str(com_lengths$Year)
```
Good! Now we are all ready to plot. 

## Plot it
```{r}
g <- ggplot(com_lengths, aes(x=Year, y=Len_cat, size=N)) + 
  geom_point(shape=21, fill="purple") + #make an empty bubble  and then fill with a color 
  scale_y_continuous(name="Length Bin", breaks = seq(lbinlow, lbinhigh, lbinbreaks))  +
  theme_bw()+
  ggtitle("Length Frequency Bubble Plot for Hogfish-Commercial Hook and Line ") 

g
```

If you want to change the size of the ramp (size of the bubbles for each level) you can do that with 
'r scale_size_continous'. 
```{r}
g + scale_size_continuous(range=c(size_range_min, size_range_max)) 
```

